---
title: "TMT Data Analysis Template"
author: "Author: Gerard Lomas | Ran By: "
date: "6/11/2022"
output: html_document
runtime: shiny
  
---
## Workflow Overview
  **Steps in Bold will require user input**

  I.  Set Working Environment 
        1. Load R Packages
        2. Create Functions
  II. **Load Initial Data Files**
        1. T_Data_Package_Analysis_Jobs 
        2. T_Reporter_Ions_Typed
        3. Mage_Processor SicStats Results
        4. Protein Collection List 
  III. Clean Data
        1. Combine T_data_Package_Analysis_Jobs with Mage_Processor_SicStats_Results
        2. **Enter Filtering Parameters (Job/Dataset_ID Selection & MSGFSpecProb Filter)**
        3. Create Unique ID (Dataset#Scan)
        4. Clean Protein Syntax
        5. Combine Initial Data into a single dataframe
  IV. Sum Redundant Peptides
  V.  Box Plots and Central Tendency Normalization 
        1. Pre-Normalization Box Plot 
        2. Central Tendency Normalization
        3. Normalized Box Plot 
        4. Anti log of Reporter Ions 
  VI. Protein Rollup 
  VII. Remove Reverse Hits and Contaminants 
  VIII. Add Protein Collection List
  IX. Statistical Analysis for each Comparision (ABP vs. NP, ABP vs. Comp)
        1. Create dataframe with only Test groups 
        2. Count Valid Values 
        3. Apply Presence/Absence Filter **(Save Resulting dataset)**
        4. Apply t-Test Valid Values Filter 
        5. Run t-test function 
        6. Run log2 Fold-Change function **(Save Resulting datast)**
  X. Create and Save Documentation and Datasets 
  
  

# **I. Set Working Environment**
```{r}
#setwd("~/Desktop/")
```

## 1. Load R Packages 
```{r, message=FALSE}
library(tidyverse)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggvis)
library(rmarkdown)
library(readxl)
library(dplyr)
library(reshape2)
library(matrixTests)
library(gtools)
library(knitr)
library(kableExtra)
library(shiny)
library(DT)

```

## 2. Create Functions
```{r}
MeanCenter.Sub <- function(Data, Mean=TRUE, centerZero=TRUE)
  {
  ndata <- Data
  
  # Compute the mean or median of each column in matrix Data, storing in vector Center
  if (Mean)
    Center <- apply(Data,2,mean,na.rm=TRUE)
  else
    Center <- apply(Data,2,median,na.rm=TRUE)
  
  # Transform vector Center into a matrix
  centerM <- matrix(Center,nrow=dim(Data)[1],ncol=dim(Data)[2], byrow=TRUE)
  
  if (centerZero)
  {
    # For data in each column, subtract the
    # column mean or median from the values, 
    # centering at zero
    ndata <- Data - centerM
  }
  else
  {
    # For data in each column, subtract the
    # column mean or median from the values, 
    # centering at the maximum mean or median
    newAverage <- max(Center,na.rm=TRUE)
    ndata <- Data - centerM + newAverage
  }
  return(ndata)
}
```


# II. Load Initial Data Files

### 1. T_Data_Package_Analysis_Jobs
```{r, message=FALSE, warning=FALSE} 
t_data_package_analysis_jobs <- read_excel(paste0(here::here(),"/Data/TMT_Practice_Dataset.xlsx"), sheet = "T_Data_Package_Analysis_Jobs")

DT::datatable(
  t_data_package_analysis_jobs,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)
```


### 2. T_reporter_Ions_Typed
```{r, message=FALSE, warning=FALSE}
T_data_df <- read_excel(paste0(here::here(),"/Data/TMT_Practice_Dataset.xlsx"), sheet = "T_Reporter_Ions_Typed")
DT::datatable(
  T_data_df,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)
```


### 3. Mage_Processor Results
```{r, message=FALSE, warning=FALSE}

mage_df <- read_excel(paste0(here::here(),"/Data/TMT_Practice_Dataset.xlsx"), sheet = "Mage")
DT::datatable(
  mage_df,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)

```


### 4. Protein Collection
```{r, message=FALSE, warning=FALSE}

Protein_collection <- read_excel(paste0(here::here(),"/Data/TMT_Practice_Dataset.xlsx"), sheet = "Sheet1")
DT::datatable(
  Protein_collection,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)
```


# III. Clean Data 

    *In these steps, combine 3 of the 4 initial datasets into one (not including Protein Collection List). Combine the Mage_Processor SicStats and T_Data_Package_Analysis_Jobs together and filter for the selected Job or Dataset_ID number. Next, set the pre-determined MSGF_SpecProb cut off. After filtering, create a unique identifier by combining the 'Dataset' and 'Scan' columns, remove astriks syntax from 'Peptide', and merge the Filtered Mage data and T_Reporter_Ions together by the unique ID ('Dataset#Scan').*

### 1. Combine T_data_Package_Analysis_Jobs with Mage_Processor_SicStats_Results
```{r, message=FALSE, warning=FALSE}
  #Merge t_data_package_analysis_jobs to mage_df by column "Job"
Mage <- merge(mage_df, t_data_package_analysis_jobs, by = ("Job"))

DT::datatable(
  Mage,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)
```


### 2. Set Filtering Parameters
***Enter/Edit the selected Mage Job for this Analysis***
```{r}
  #Select certain job ID 
Mage_job = Mage %>% filter(Job == 2044576)

```


***Enter/edit the MSGf_SpecProb value (for rat, set to <= 2.7113E-8)***
```{r}
#Filter for pre-set FDR (DTRA Rat data MSGF_SpecProb <= 2.7113E-8) ###This is calculated based on an analysis of the global proteomics of the test organism
Mage_fdr_df = Mage_job %>% filter(MSGF_SpecProb <= 2.7113E-8)

```


###### Remove unnecessary columns from Mage_fdr_df
```{r, message=FALSE, warning=FALSE}
  #3. Remove unnecessary columns from Mage_fdr_df -> Create new Mage_df with only 'Job', 'QValue', 'MSGF_SpecProb', 'Peptide', 'Protein', 'Scan' columns
Mage_fdr_filtered = Mage_fdr_df[, c('Job', 'Dataset_ID', 'QValue', 'MSGF_SpecProb', 'Peptide', 'Protein', 'Scan')]


```


### 3. Create Unique Identifiers (Dataset#Scan)
```{r, message=FALSE, warning=FALSE}
#Create and Merge Dataset#Scan Columns for Mage_fdr_filtered and T_data_df ####
  #Create Dataset#Scan Column for df: T_data_df 
Mage_fdr_filtered$DatasetScan = paste(Mage_fdr_filtered$Dataset_ID, '#', Mage_fdr_filtered$Scan)

  #Create Dataset#Scan Column for df: T_data_df 
T_data_df$DatasetScan = paste(T_data_df$Dataset, '#', T_data_df$ScanNumber)

  #Inner_join (merge) Mage_fdr_filtered and T-data_df into new df (merged_df) by column DatasetScan for all Mage_fdr_filtered entries
merged_df = inner_join(Mage_fdr_filtered, T_data_df, by = "DatasetScan")

```


### 4. Clean Peptide Syntax
```{r, message=FALSE, warning=FALSE}
  #Remove "*" from peptides and replace with "" (gsub function -> Removes 3+ occurances)
merged_df$Peptide <- gsub("*","",as.character(merged_df$Peptide), fixed = TRUE)

  #Check for and count remaining "*" in peptide 
merged_df$Astrisk_count <- str_count(merged_df$Peptide,'\\*')
#Verify all * have been removed by checking Astrisk_count column. All should have 0 value
```

### 5. Combine Initial Data Into a Single Dataframe
**Check for both dataset, scan, etc. columns.
```{r}

  #3. Create new df with only 'Job', 'QValue', 'Peptide', 'Protein', 'Ion Values' columns
Combined_df = merged_df[, c( 1, 3, 5:6, 16:25)]

DT::datatable(
  Combined_df,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)

#**********************Display Columns ***********************************
```


# IV. Sum Redundant Peptides

**Combine Redundant Peptides and sum their reporter ion values. Apply a counter to determine the amount of redundancy per peptide**
```{r, message=FALSE, warning=FALSE}
# SUM REDUNTANT REPORTER ION INTENSITIES ####
# Remove peptide redundancy and sum Reporter Ion values

# This line will combine duplicate Peptides and sum all ions
    # detach("package:plyr", unload = TRUE) ###This line is needed if dplyr error n()) occurs 
Pep_redundency <- Combined_df %>%
  group_by(Peptide) %>%
  summarise(across(Ion_126.128:Ion_131.138, sum))

# Create df with columns that were removed in Pep_redundency
  # Counts the number of duplicate $Peptides
Missing_Columns <- Combined_df %>%
  group_by(Peptide) %>% 
  mutate(Pep_count = n()) 
  # Remove all reporter Ion columns 
Missing_Columns <- Missing_Columns %>%
  select('Job','QValue', 'Protein', 'Peptide' , 'Pep_count')
  # Remove all redundancy of $Peptide and other columns missing from df1 
Missing_Columns <- distinct(Missing_Columns, Peptide, .keep_all = TRUE)

# Merge these two dataframes together and (B)re-organize 
Analysis_df = merge(Missing_Columns, Pep_redundency, by = 'Peptide', all.x = FALSE) 
# Re-organize the df 
Analysis_df = Analysis_df %>%
  select('Job', 'QValue', 'Protein', 'Peptide', 'Pep_count', 6:15)

```

**Analysis dataframe - This datatable should have a column for Job, QValue, Protein, Peptide, Pep_count, and all samples (reference sample, ABP, Comp, NP)**
```{r}
DT::datatable(
  Analysis_df,
  clas='hover cell-border stripe',
  options = list(scrollX = TRUE, pageLength=5)
)

```


# V. Box Plots and Central Tendency Normalization
### 1. Pre-Normalization Box Plot 
```{r, message=FALSE, warning=FALSE}
# LOG TRANSFORMATION, CENTRAL TENDENCY NORMALIZATION, ANTILOG ####
# Create Box blot before Central Tendency Normalization
  # Group samples for box plot (reference, ABP, Competitor, NP)
bp_initial = Analysis_df %>%
  select( 6:15)
colnames(bp_initial) <-c('ref', 'ABP1','ABP2', 'ABP3', 'Comp1', 'Comp2', 'Comp3', 'NP1', 'NP2', 'NP3')

bp_initial = log2(bp_initial)
  # melt "test" into Variables and values for boxplot
bp_initial_plot <- melt(bp_initial)

  # Initial Box Plot 
ggplot(bp_initial_plot, aes(
  x = variable, 
  y = value, 
  fill = variable), na.rm=TRUE) +
  geom_boxplot() 


```


### 2. Central Tendency Normalization

**Central Tendency Normalization is applied to each sample group separately**

```{r, message=FALSE, warning=FALSE}

# Perform log2 Transformation (reduces varience)
log2_data <- Analysis_df
log2_data[,6:15] <- log(log2_data[6:15], 2)

# Separate sample groups by creating individual dataframes
  # df with only ABP samples 
Cen_ten_ABP <- log2_data[7:9]
  # Remove infinite errors 
is.na(Cen_ten_ABP) <- do.call(cbind, lapply(Cen_ten_ABP, is.infinite)) 
  # df with only Comp samples
Cen_ten_Comp <- log2_data[10:12]
  # Remove infinite errors 
is.na(Cen_ten_Comp) <- do.call(cbind, lapply(Cen_ten_Comp, is.infinite))
  # df with only NP samples
Cen_ten_NP <- log2_data[13:15]
  # Remove infinite errors 
is.na(Cen_ten_NP) <- do.call(cbind, lapply(Cen_ten_NP, is.infinite))

#Apply Central Tendency Normalization for each sample group (Probe, competitor and NP)
      #Function workflow -> (1) Compute mean of each column/replicate. (2) subtract the column mean from the values while centering at the max mean)
Mean_cen_ABP <- MeanCenter.Sub(Cen_ten_ABP, Mean=TRUE, centerZero = FALSE)
Mean_cen_Comp <- MeanCenter.Sub(Cen_ten_Comp, Mean=TRUE, centerZero = FALSE)
Mean_cen_NP <- MeanCenter.Sub(Cen_ten_NP, Mean=TRUE, centerZero = FALSE)

# Combine results into one df 
Normalized_df2 <- cbind(log2_data[1:6], Mean_cen_ABP, Mean_cen_Comp, Mean_cen_NP)
```

### 3. Normalized Box Plot
```{r, message=FALSE, warning=FALSE}

# Create Box Plot of the Normalized data
bp_normal = Normalized_df2 %>%
  select(7:15) 

bp_normal_plot <- melt(bp_normal)

# Box Plot 
ggplot(bp_normal_plot, aes(
  x = variable, 
  y = value, 
  fill = variable), na.rm=TRUE) +
  geom_boxplot() 

```

### 4. Antilog of Reporter Ion Values 
```{r,  message=FALSE, warning=FALSE}
# Antilog of reporter ions 
  # Copy Normalized_df2 dataframe
Antilog_df <- Normalized_df2
  # Remove infinate errors (NA) from reporter ion column
is.na(Antilog_df$Ion_126.128) <- do.call(cbind, lapply(Antilog_df$Ion_126.128, is.infinite))
  # Take antilog of all reporter ion columns 
Antilog_df[,6:15] <-  2^(Antilog_df[6:15])

```


# VI. Protein Roll Up 
**Combine Redundant Proteins and sum their reporter ion values. Apply a counter (Unique_Pep_count) to determine the amount of Unique peptides per Protein. Sum the redundant peptides to get total peptides per protein (Pep_count)**
```{r, message=FALSE, warning=FALSE}
# PROTEIN ROLLUP ####
# Protein Rollup - Combine duplicate Proteins and sum all ions
  # Duplicate Antilog_df 
Pro_redundency <- Antilog_df
  # Combine duplicate Proteins and sum all ions
Pro_redundency <- Pro_redundency%>%
  group_by(Protein) %>%
  summarise(across(Ion_126.128:Ion_131.138,sum, na.rm = TRUE))
  # Change 0 values back to NA
Pro_redundency <- na_if(Pro_redundency, 0)

# Create df with columns that were removed in Antilog_df
  # Counts the number of duplicate $Proteins into Unique_pep_count
Missing_Columns_pro <- Antilog_df %>%
  group_by(Protein) %>% 
  mutate(Unique_Pep_count = n()) 

  # Remove all reporter Ion columns 
Missing_Columns_pro <- Missing_Columns_pro %>%
  select('Job','QValue', 'Protein', 'Unique_Pep_count' , 'Peptide' , 'Pep_count')

  # Create new df with column with total protein count
Total_pep_count <- Missing_Columns_pro %>%
  group_by(Protein) %>%
  summarise(across(Pep_count, sum))

  # Important - this will remove all $Protein redundancy 
Missing_Columns_pro2 <- distinct(Missing_Columns_pro, Protein, .keep_all = TRUE)
  # Create C
Missing_Columns_pro2 <- Missing_Columns_pro2[1:5]

#3. Merge the two dataframes together and (B)re-organize 
  # Merge Missing, Col, with Total_pep_count 
Roll_up_df = merge(Missing_Columns_pro2, Total_pep_count,  by = 'Protein', all.x = FALSE) 
  # Merge with Pro_redundency 
Roll_up_df2 = merge(Roll_up_df, Pro_redundency, by = 'Protein', all.x = FALSE) 
  # Remove 'Peptide' and 'QValue' Columns 
Roll_up_df2 <- subset(Roll_up_df2, select = -Peptide)
Roll_up_df2 <- subset(Roll_up_df2, select = -QValue)

  

DT::datatable(
  Roll_up_df2,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=10, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```



# VII. Remove Reverse Hits and Contaminants
**Create two columns to ID rows as TRUE/FALSE if Protein column contains "XXX" or "Contaminant." Remove all rows that contain a TRUE conditional in column "XXX" or "Contaminant"**
```{r}
# Add observation counts 
  # Copy Roll_up_df2 dataframe  
xxx_rm <- Roll_up_df2 

  # Create column "XXX" and "Contaminants and ID if Protein column contains these elements 
xxx_rm$XXX = grepl("XXX", xxx_rm$Protein, ignore.case = TRUE)
xxx_rm$Contaminant = grepl("Contaminant", xxx_rm$Protein, ignore.case = TRUE)

  # Create dataframe Stat_df containing rows where XXX == FALSE and Contaminant == FALSE elements
Stat_df = filter(xxx_rm, xxx_rm$XXX == FALSE & xxx_rm$Contaminant == FALSE)
```

**Scroll to the right to check if all XXX and Reverse Hits have been removed. All values should be 'FALSE'**
```{r}
  # Display Stat_df Dataframe
DT::datatable(
  Stat_df,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=10, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```

# VIII. Add Protein Collection List 
```{r}

Stat_Protein_df <- merge(x = Stat_df, y = Protein_collection, by.x = ("Protein"), by.y = ("Protein_Name"), all.x = TRUE)


DT::datatable(
  Stat_Protein_df,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```


# IX. Statistical Analysis for each Comparision (ABP vs. NP)
### 1. Create Dataframe with Only Test Group Reporter Ions (ABP, NP)
  **Check to make sure ABP and NP columns are in this dataframe**

```{r}
ABP_NP_df <- cbind(Stat_Protein_df[1:4], Stat_Protein_df[6:8], Stat_Protein_df[12:14], Stat_Protein_df[17:23])

DT::datatable(
  ABP_NP_df,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```


### 2. Count Valid Values (ABP vs. NP)
```{r}
# Count for ABP samples per protein
ABP_NP_df$Count_ABP <- rowSums(!is.na(ABP_NP_df[,5:7]))

# Count for NP samples per protein 
ABP_NP_df$Count_NP <- rowSums(!is.na(ABP_NP_df[,8:10]))

```

### 3. Apply Presence Absence Filter (ABP vs. NP)

**Save Resulting Dataset (ABP vs. NP Pres_Abs Results) **
```{r}
ABP_NP_Pres_Abs = filter(ABP_NP_df, ABP_NP_df$Count_ABP >= 2 & ABP_NP_df$Count_NP <= 1)

DT::datatable(
  ABP_NP_Pres_Abs,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
) 
```

### 4. Apply t-Test Valid Values Filter (ABP vs. NP)
```{r}
ABP_NP_ttest = filter(ABP_NP_df, ABP_NP_df$Count_ABP >= 2 & ABP_NP_df$Count_NP > 1)

```

### 5. Run t-Test Function (ABP vs. NP)

```{r}
# Log2 Transform Reporter Ions 
ABP_NP_log2 <- cbind(ABP_NP_ttest[1:4], log2(ABP_NP_ttest[5:10]), ABP_NP_ttest[11:19]) 

# t-Test function ABP vs. NP
ABP_NP_log2_Final <- cbind(ABP_NP_log2, row_t_welch(ABP_NP_log2[5:7], ABP_NP_log2[8:10], alternative = "greater"))

```

### 6. Run log2 Fold-Change Function (ABP vs. NP)
```{r}
# Re-organize dataframe to add t-test Results 
ABP_NP_Results <- ABP_NP_log2_Final

# Calculate and append Fold Change
ABP_NP_Results$Log2FC <- (rowMeans(ABP_NP_log2_Final[5:7], na.rm = TRUE) - rowMeans(ABP_NP_log2_Final[8:10], na.rm = TRUE))
```
### ABP vs. NP t_test and log2 FC Results
**Save Resulting Dataset (ABP vs. NP t-Test Results)**
```{r}
DT::datatable(
  ABP_NP_Results,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)
```

**Re-Orangize Dataframe**
```{r}
# Note *Consider appending base and log2 df by Protein and not cbind*
#ABP_NP_ttest_results <- cbind(ABP_NP_Results[1:2], ABP_NP_ttest[3:8], ABP_NP_Results[3:8], ABP_NP_Results[10], ABP_NP_Results[9])

```


# IX. Statistical Analysis for each Comparison (ABP vs. Comp)
### 1. Create Dataframe with Only Test Group Reporter Ions (ABP, Comp)
  **Check to make sure ABP and Comp Reporter Ion columns are in this dataframe**

```{r}
ABP_Comp_df <- cbind(Stat_Protein_df[1:4], Stat_Protein_df[6:8], Stat_Protein_df[9:11], Stat_Protein_df[17:23])

DT::datatable(
  ABP_Comp_df,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```


### 2. Count Valid Values (ABP vs. Comp)
```{r}
# Count for ABP samples per protein
ABP_Comp_df$Count_ABP <- rowSums(!is.na(ABP_Comp_df[,5:7]))

# Count for Comp samples per protein 
ABP_Comp_df$Count_Comp <- rowSums(!is.na(ABP_Comp_df[,8:10]))

```

### 3. Apply Presence Absence Filter (ABP vs. Comp)

**Save Resulting Dataset (ABP vs. Comp Pres_Abs Results) **
```{r}
ABP_Comp_Pres_Abs = filter(ABP_Comp_df, ABP_Comp_df$Count_ABP >= 2 & ABP_Comp_df$Count_Comp <= 1)

DT::datatable(
  ABP_Comp_Pres_Abs,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)

```

### 4. Apply t-Test Valid Values Filter (ABP vs. Comp)
```{r}
ABP_Comp_ttest = filter(ABP_Comp_df, ABP_Comp_df$Count_ABP >= 2 & ABP_Comp_df$Count_Comp > 1)

```

### 5. Run t-Test Function (ABP vs. NP)
```{r}
# Log2 Transform Reporter Ions 
ABP_Comp_log2 <- cbind(ABP_Comp_ttest[1:4], log2(ABP_Comp_ttest[5:10]), ABP_Comp_ttest[11:19]) 

# t-Test function ABP vs. Comp
ABP_Comp_log2_Final <- cbind(ABP_Comp_log2, row_t_welch(ABP_Comp_log2[5:7], ABP_Comp_log2[8:10], alternative = "greater"))
```


### 6. Run log2 Fold-Change Function (ABP vs. Comp)
```{r}
# Re-organize dataframe to add t-test Results 
ABP_Comp_Results <- ABP_Comp_log2_Final

# Calculate and append Fold Change
ABP_Comp_Results$Log2FC <- (rowMeans(ABP_Comp_log2_Final[5:7], na.rm = TRUE) - rowMeans(ABP_Comp_log2_Final[8:10], na.rm = TRUE))
```

### ABP vs. Comp Results
**Save Resulting Dataset (ABP vs. Comp t-Test Results)**
```{r}
DT::datatable(
  ABP_Comp_Results,
  clas='hover cell-border stripe',
  extensions = 'Buttons',
  options = list(scrollX = TRUE, 
                 pageLength=5, 
                 dom='Bfrtip',
                 buttons = c('copy', 'csv', 'excel', 'pdf', 'print'))
)
```


```{r}
# Note *Consider appending base and log2 df by Protein and not cbind*
#ABP_Comp_ttest_results <- cbind(ABP_Comp_Results[1:2], ABP_Comp_ttest[3:8], ABP_Comp_Results[3:8], ABP_Comp_Results[10], ABP_Comp_Results[9])

```

# X. Create and Save Documentation and Datasets 
**Save the code file through R.studio. (2) Save datasets from user interface. (3) Save user interface rMarkdown Document by selecting "Open in Browser" tab (top left), then save as pdf. 


###Session Info: 
```{r}
sessionInfo()
```